Embedding Intuitionistic into Classical Logic

### 2022-07-16 CSL/lipics-v2021-EIICL.tex
proof read of whole paper (without appendix) finished

Fix some set $\mathcal S$ of formulas of the form $$A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C), (A\vee B)\to C, (A\to B)\to C$$and$$\varphi = \bigwedge \mathcal S\to P,$$
and denote with $\mathcal F_\to$ the set of formulas of the form $(A\to B)\to C$. We will now formulate a QBF that expresses that $\varphi$ has an intuitionistic counter-model. The idea how to get a linear-size formula is: Instead of expressing validity at every node of $\mathcal M_T$ - which was done in the CPC translation - we express that on each path in $\mathcal M_T$ the nodes satisfy all conditions for $\mathcal M_T$ being a counter-model. The universally quantified variables $F_\psi$ in the next definition handle the branching, i.e. express which path is considered.

\begin{definition}
	For every propositional variable $X$ occurring in $\mathcal S$ and non-negative integer $n$ consider a new atom $X^n$ and for every formula $\psi\in\mathcal F_\to$ consider a new atom $F_\psi^n$. Let $\vec X^n$ range over all propositional variables $X^n$ and $\vec F^n$ range over all $F_\psi^n$. Define
	\begin{itemize}
		\item {\normalfont Valid}$(n)$ which encodes that $\vec F^n_\psi$ represents a valid next element of a sequence without repetition, i.e. if we interpret the $\vec F^i$ as bit-vectors, then $\vec F^n$ has exactly one bit set to $1$, indicating the formula at the $n$-th position of the sequence, and it is at a position that is $0$ in $\sum\{\vec F^i\:|\:i < n\}$.
		\item {\normalfont Persistent}$(n)$ which encodes that the persistency condition holds.
		\item {\normalfont Sat}$_{\mathcal S}(n)$ encoding that the formulas in $\mathcal S\setminus\mathcal F_\to$ hold at the $n$-th world of the path.
		\item {\normalfont Sat}$_{\mathcal F_\to}(n)$ encoding that the formula in $\mathcal F_\to$ that is represented by $\vec F^{n-1}$ holds.
	\end{itemize}
	Have $k = |\mathcal F_\to|$ and define$$\varphi^Q_{i} = \exists \vec X^i\forall \vec F^i\left(\text{\normalfont Persistent}(i)\wedge \text{\normalfont Sat}_{\mathcal S}(i)\wedge \text{\normalfont Sat}_{\mathcal F_\to}(i)\wedge\left(\text{\normalfont Valid}(i)\to \varphi^Q_{i+1}\right)\right)$$
	for $0 < i < k$ as well as the special cases
	$$\varphi^Q_{k} = \exists \vec X^k\left(\text{\normalfont Persistent}(k)\wedge \text{\normalfont Sat}_{\mathcal S}(k)\wedge \text{\normalfont Sat}_{\mathcal F_\to}(k)\right)$$
	for leafs and $$\varphi^Q = \varphi_0^Q = \exists \vec X^0\forall \vec F^0\left(\neg P^0\wedge \text{\normalfont Sat}_{\mathcal S}(0)\wedge(\text{\normalfont Valid}(0)\to \varphi_1^Q)\right)$$for the root.
	Example encodings of the above formulas would be
	\begin{align*}
		\text{\normalfont Valid}(n) = &\left(\bigvee \{\mathcal F^n_\psi\:|\:\psi\in\mathcal F_\to\}\right)\wedge\left(\bigwedge \{\neg(F^n_{\psi_1}\wedge F^n_{\psi_1})\:|\:\psi_1\neq\psi_2\in\mathcal F_\to\}\right)\wedge\\&\left(\bigwedge\left\{\bigwedge\{F^i_\psi\to\neg F^n_\psi\:|\:i < n\}\:|\:\psi\in\mathcal F_\to\right\}\right)\\
		\text{\normalfont Persistent}(n) = & \bigwedge\{X^{n-1}\to X^n\:|\: X \text{ prop. variable with }X^{n-1}\in\vec X^{n-1}, X^n \in\vec X^{n}\}\\
		\text{\normalfont Sat}_{\mathcal S}(n) = &\bigwedge\left\{(A^n\circ B^n)\to C^n\:|\:(A\circ B)\to C\in \mathcal S\setminus\mathcal F_\to\right\}\wedge\\
		&\bigwedge\left\{A^n\to (B^n\circ C^n)\:|\:A\to (B\circ C)\in \mathcal S\setminus\mathcal F_\to\right\}\\
		\text{\normalfont Sat}_{\mathcal F_\to}(n) = & \bigwedge\{F^{n-1}_\psi\to (A^n\to B^n)\to C^{n-1}\:|\:\psi = (A\to B)\to C\in\mathcal F_\to\}
	\end{align*}
\end{definition}

\begin{example}
	With the above encodings for double negation elimination $\varphi = ((A\to \bot)\to \bot)\to A$ we have
	$$\hspace*{-.3cm}
	\varphi^Q = \exists A^0\forall F^0\left(\neg A^0\wedge\left(F^0\to \exists A^1\left((A^0\to A^1)\wedge(F_0\to (A^1\to \bot)\to \bot)\right)\right)\right)
	$$
	which is a satisfiable QBF since $A^0 = 0, A^1 = 1$ satisfies it for any $F^0$.
\end{example}

\begin{lemma}
	$\varphi$ is not intuitionistically valid iff $\varphi^Q$ is a satisfiable QBF.
\end{lemma}
A proof can be found in the appendix as proof~\ref{lemma:QBF}.

A simple counting argument shows that for each $n < |\mathcal F_\to|$ we have $|\text{Valid}(n)|\in \mathcal O(|\mathcal F_\to|^2)$, $|\text{Persistent}(n)|, |\text{Sat}_{\mathcal S}(n)|\in \mathcal O(|\varphi|)$ and $|\text{Sat}_{\mathcal F_\to}(n)|\in \mathcal O(|\mathcal F_\to|)$ so overall we get
\begin{lemma}
	The size of $\varphi^Q$ is in $\mathcal O(|\varphi|\cdot|\mathcal F_\to| + |\mathcal F_\to|^3)$.
\end{lemma}

@@

First of all recall that due to Lemma~\ref{equivalid} it is sufficient to consider sentences of the form $\bigwedge\mathcal S\to P$ where all formulas in $\mathcal S$ are of one of the forms
\[\begin{matrix}
	\forall \vec z(A(\vec a)\to (B(\vec b)\wedge C(\vec c)))\indent \forall \vec z((A(\vec a)\wedge B(\vec b))\to C(\vec c))\indent \forall \vec z(A(\vec a)\to (B(\vec b)\vee C(\vec c)))\\
	\forall \vec z((A(\vec a)\vee B(\vec b))\to C(\vec c))\indent \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c))\indent\forall \vec z(\forall xA(\vec a)\to B(\vec b))\\ \forall \vec z(A(\vec a)\to\forall xB(b)), \forall \vec z(\exists xA(\vec a)\to B(\vec b)), \forall \vec z(A(\vec a)\to\exists xB(b))
\end{matrix}\]
The complete result that we will show is as follows:
\begin{theorem}\label{fullFOtranslation}
	Let $\mathcal S$ be as above and $\mathcal F_\to\subseteq\mathcal S$ contain the formulas of the form $\forall \vec t((A(\vec z_A)\to B(\vec z_B))\to C(\vec z_C))$ and $\mathcal F_\forall\subseteq\mathcal S$ the formulas of the form $\forall \vec z(\forall xA(\vec z_A)\to B(\vec z_B))$. Let $\mathcal F = \mathcal F_\to\cup\mathcal F_\forall$. For every $n$-ary atomic relation $R$ define a new $n+1$-ary relation $R^\#$. For every $\psi\in\mathcal F$ define a new function symbol $f_\psi$. Consider a new binary relation symbol $E$ and define $\vec E(\vec x, u) := \bigwedge\{E(x, u)\:|\:x\in\vec x\}$. Obtain $\mathcal S^\#$ by including formulas as follows:
	\begin{itemize}
		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, u)\to (B^\#(\vec b, u)\circ C^\#(\vec c, u))))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z(A(\vec a)\to (B(\vec b)\circ C(\vec c)))\in\mathcal S$
		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, u)\circ B^\#(\vec b, u))\to C^\#(\vec c, u))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z((A(\vec a)\circ B(\vec b))\to C(\vec c))\in\mathcal S$
		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to(A^\#(\vec a, f(\vec z, u))\to B^\#(\vec b, f(\vec z, u)))\to C^\#(\vec c, u))$ for each $\psi = \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c))\in\mathcal S$
		\item  $\forall \vec z\forall u(\vec E(\vec z, u)\to \forall x(E(x, f(\vec z, u))\to A^\#(\vec a, f(\vec z, u)))\to B^\#(\vec b, u))$\\for each $\forall \vec z(\forall xA(\vec a)\to B(\vec t_B))\in\mathcal S$.
		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to A^\#(\vec a, u)\to \forall x(E(x, u)\to B^\#(\vec b, u)))$\\for each $\forall \vec t(A(\vec a)\to \forall xB(\vec b))\in\mathcal S$.
		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to \exists x(E(x, u)\wedge A^\#(\vec z, u))\to B^\#(\vec b))$\\for each $\forall \vec z(\exists xA(\vec a)\to B(\vec b))\in\mathcal S$
		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to A^\#(\vec b)\to \exists x(E(x, u)\wedge B^\#(\vec b, u)))$\\for each $\forall \vec z(A(\vec a)\to \exists xB(\vec b))\in\mathcal S$
	\end{itemize}
	Then $\bigwedge S\to P$ is intuitionistically iff $\exists xE(x, s)\to K\to \bigwedge\mathcal S^\#\to P^\#(s)$ is classically valid, where $s$ is a new constant symbol, $\Sigma$ is the original signature and
	\begin{align*}
		K := &\:\bigwedge\{\forall x\forall u((E(x, u)\to E(x, f_\psi(u)))\:|\:\psi\in\mathcal F\}\wedge\\
		&\:\forall u\left(\exists xE(x, u)\to \bigwedge\{\forall\vec z(\vec E(\vec z, u)\to E(f(\vec z), u))\:|\:\text{$f$ is a function symbol in $\Sigma$}\}\right)\wedge\\
		&\:\bigwedge\{\forall\vec z_1\forall \vec z_2\forall u(A^\#(\vec z_1, u)\to A^\#(\vec z_1, f_\psi(\vec z_2, u)))\:|\:\text{$\psi\in\mathcal F$, $A$ is a relation symbol in $\Sigma$}\}
	\end{align*}
	The size of the obtained formula is linear in the input. However a $|\mathcal F| + 1$ new function symbols and the new binary predicate $E$ were introduced and each $n$-ary relation symbol has been extended to a $n+1$-ary relation symbol.
\end{theorem}

It is not possible to completely eliminate worlds from the domain as in the propositional case, since counter-models possibly must be infinite. We will see why that is in more detail later.

@@

As in the propositional case we must encode the Kripke Semantics, this time not only the relation $\preceq$ but also $E$ has to be axiomatized. For that we use the following predicates:
\begin{itemize}
	\item PartialOrder($\preceq$) encoding that $\preceq$ is a partial order.
	\item DomainSubset$(u, w)$ encoding that the domain at $u$ is a subset of that at $w$.
	\item World($u$) encoding that $u$ represents a world.
	\item DomainClosed($u$) which encodes that at $u$ the domain is closed under functions, e.g. contains all interpretations of constants.
	\item Persistent$(u, w)$ which encodes that persistency is satisfied between $u$ and $w$, i.e. for all predicates $A$ and domain elements $x$ with $A(x)$ at $u$ we have $A(x)$ at $w$.
\end{itemize}
Then analogously to the predicate case define
\begin{align*}
	K(\varphi) = \:& \text{PartialOrder}(\preceq) \wedge \forall u \forall w (u\preceq w\to \text{DomainSubset}(u, w)) \wedge\\
				 & \forall u(\text{World(u)}\to \text{DomainClosed}(u))\wedge \forall u\forall w (u\preceq w\to \text{Persistent}(u, w))
\end{align*}
Example encodings could be
\begin{align*}
	\text{PartialOrder}(\preceq) = &\:\forall u(u\preceq u)\wedge\forall u\forall w(u\preceq w\to w\preceq u\to u = w)\wedge\\&\:\forall u\forall v\forall w(u\preceq v\to v\preceq w\to u\preceq w)\\
	\text{DomainSubset}(u, w) = &\:\forall z(E(z, u)\to E(z, w))\\
	\text{World}(u) = &\:\exists xE(x, u)\\
	\text{DomainClosed}(u) = &\:\bigwedge\{\forall\vec z(\vec E(\vec z, u)\to E(f(\vec z), u))\:|\:\text{$f\in\Sigma$ is a function symbol}\}\\
	\text{Pesistent}(u, w) = &\:\bigwedge\{\forall\vec z(\vec A^\#(\vec z, u)\to A^\#(\vec z, w))\:|\:\text{$A\in\Sigma$ is a relation symbol}\}
\end{align*}

@@

Then $\varphi = \bigwedge\mathcal S\to P$ is intuitionistically valid if and only if
$$\varphi^\circ= \text{World}(s)\to K(\varphi)\to \bigwedge S^\circ\to P^\#(s)$$
is classically valid where $s$ is a new constant symbol. This was obvious in the propositional case but here it is more nuanced. A proof is given in the appendix as Lemma~\ref{lemma:fo-simplification}.

@@

As in the predicate case we now show the existence of a canonical counter-model, if a counter-model exists. Denote the set of formulas $\psi\in\mathcal S$ of the form $\forall\vec z((A(\vec a)\to B(\vec b))\to C(\vec c))$ with $\mathcal F_\to$ and of the form $\forall \vec z(\forall xA(\vec a)\to B(\vec b))$ with $\mathcal F_\forall$. Let $\mathcal F = \mathcal F_\to\cup\mathcal F_\forall$. For each  $\psi\in\mathcal F$ consider a new function symbol $f_\psi$. As before we may assume that $s$ is a least element and eliminate the quantification over $w$ by adding formulas
\begin{itemize}
	\item $\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, f_\psi(\vec z, u))\to B^\#(\vec b, f_\psi(\vec z, u)))\to C^\#(\vec c, u))$\\instead of $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to A^\#(\vec a, w)\to B^\#(\vec b, w))\to C^\#(\vec c, u))$
	\item $\forall \vec z\forall u(\vec E(\vec z, u)\to (\forall x(E(x, f_\psi(\vec z, u))\to A^\#(\vec a, f_\psi(\vec z, u))))\to B^\#(\vec b, u))$\\instead of $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to \forall x(E(x, w)\to A^\#(\vec a, w)))\to B^\#(\vec b, u))$
\end{itemize}
and in each other formula in $\mathcal S^\circ$ removing the $s\preceq u$ to obtain $\mathcal S^\#$ from $\mathcal S^\circ$. Then the previous formulas are equivalid to
$$\varphi^\circ= \forall u(s\preceq u)\to \bigwedge\{\forall \vec t\forall u(u\preceq f(\vec z, u))\:|\:\psi\in\mathcal F\}\to K(\varphi)\to\bigwedge\mathcal S^\#\to P^\#(s)$$
As in the previous section we can define a tree counter-model for this.

\begin{definition}
	Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^\circ$. Define $\mathcal M_T = (M_T, I_T)$ as follows:
	\begin{itemize}
		\item $M_T$ are the sequences on $\{(\vec z, \psi)\:|\:\psi\in \mathcal F, \vec z\in M^n\}$.
		\item Interpret $\preceq$ as the prefix-order.
		\item Have $$f_\psi^{I_T}(\vec z, (\vec z_1, \psi_1)\dots (\vec z_n, \psi_n)) = (\vec z_1, \psi_1)\dots (\vec z_n, \psi_n)(\vec z, \psi)$$
		\item For propositional variables $P$ have $$P^{I_T}((\vec z_1, \psi_1)\dots (\vec z_n, \psi_n)) = P^I(f_{\psi_n}(\vec z_n, \dots(f_{\psi_1}(\vec z_1, s^I))\dots))$$
	\end{itemize}
\end{definition}

\begin{lemma}
	If $\mathcal M$ is a counter-model to $\varphi^\circ$ then so is $\mathcal M_T$.
\end{lemma}

Even if we don't have the nice finiteness of the propositional case we have still reduced the model complexity in that we know that considering worlds of the form $$f_{\psi_1}(\vec z_1, f_{\psi_2}(\vec z_2, \dots f_{\psi_n}(\vec z_n, s)\dots))$$ is sufficient. In particular this allows us to completely eliminate $\preceq$ from the formula, that is we remove the PartialOrder$(\preceq)$
from $K(\varphi)$ and replace $\forall u\forall w\forall x((E(x, u)\wedge u\preceq w)\to E(x, w))$ with $\bigwedge\{\forall \vec z\forall u(E(x, u)\to E(x, f_\psi(\vec z, u)))\:|\:\psi\in\mathcal F_\to\}$ and analogously for persistency to obtain $K$. Define

>>>>

Fix some set $\mathcal S$ of formulas of the form $$A\to (B\wedge C), (A\wedge B)\to C, A\to (B\vee C), (A\vee B)\to C, (A\to B)\to C$$and$$\varphi = \bigwedge \mathcal S\to P,$$
and denote with $\mathcal F_\to$ the set of formulas of the form $(A\to B)\to C$. We will now formulate a QBF that expresses that $\varphi$ has an intuitionistic counter-model. The idea how to get a polynomially-sized formula is: Instead of expressing validity at every node of $\mathcal M_T$ --- which was done in the CPC translation --- we express that on each path in $\mathcal M_T$ the nodes satisfy all conditions for $\mathcal M_T$ being a counter-model. The universally quantified variables $F_\psi$ in the next definition handle the branching, i.e. express which path is considered.

\begin{definition}
	For every propositional variable $X$ occurring in $\mathcal S$ and non-negative integer $n$ consider a new atom $X^n$ and for every formula $\psi\in\mathcal F_\to$ consider a new atom $F_\psi^n$. Let $\vec X^n$ range over all propositional variables $X^n$ and $\vec F^n$ range over all $F_\psi^n$. Define
	\begin{itemize}
		\item {\normalfont Valid}$(n)$ which encodes that $\vec F^n_\psi$ represents a valid next element of a sequence without repetition, i.e. if we interpret the $\vec F^i$ as bit-vectors, then $\vec F^n$ has exactly one bit set to $1$, indicating the formula at the $n$-th position of the sequence, and it is at a position that is $0$ in $\sum\{\vec F^i\:|\:i < n\}$.
		\item {\normalfont Persistent}$(n)$ which encodes that the persistency condition holds.
		\item {\normalfont Sat}$_{\mathcal S}(n)$ encoding that the formulas in $\mathcal S\setminus\mathcal F_\to$ hold at the $n$-th world of the path.
		\item {\normalfont Sat}$_{\mathcal F_\to}(n)$ encoding that the formula in $\mathcal F_\to$ that is represented by $\vec F^{n-1}$ holds.
	\end{itemize}
	Have $k = |\mathcal F_\to|$ and define$$\varphi^Q_{i} = \exists \vec X^i\forall \vec F^i\left(\text{\normalfont Persistent}(i)\wedge \text{\normalfont Sat}_{\mathcal S}(i)\wedge \text{\normalfont Sat}_{\mathcal F_\to}(i)\wedge\left(\text{\normalfont Valid}(i)\to \varphi^Q_{i+1}\right)\right)$$
	for $0 < i < k$, as well as the special cases
	$$\varphi^Q_{k} = \exists \vec X^k\left(\text{\normalfont Persistent}(k)\wedge \text{\normalfont Sat}_{\mathcal S}(k)\wedge \text{\normalfont Sat}_{\mathcal F_\to}(k)\right)$$
	for leafs, and $$\varphi^Q = \varphi_0^Q = \exists \vec X^0\forall \vec F^0\left(\neg P^0\wedge \text{\normalfont Sat}_{\mathcal S}(0)\wedge(\text{\normalfont Valid}(0)\to \varphi_1^Q)\right)$$for the root.
	Example encodings of the above formulas are:
	\begin{align*}
		\text{\normalfont Valid}(n) = &\left(\bigvee \{\mathcal F^n_\psi\:|\:\psi\in\mathcal F_\to\}\right)\wedge\left(\bigwedge \{\neg(F^n_{\psi_1}\wedge F^n_{\psi_1})\:|\:\psi_1\neq\psi_2\in\mathcal F_\to\}\right)\wedge\\&\left(\bigwedge\left\{\bigwedge\{F^i_\psi\to\neg F^n_\psi\:|\:i < n\}\:|\:\psi\in\mathcal F_\to\right\}\right)\\
		\text{\normalfont Persistent}(n) = & \bigwedge\{X^{n-1}\to X^n\:|\: X \text{ prop. variable with }X^{n-1}\in\vec X^{n-1}, X^n \in\vec X^{n}\}\\
		\text{\normalfont Sat}_{\mathcal S}(n) = &\bigwedge\left\{(A^n\circ B^n)\to C^n\:|\:(A\circ B)\to C\in \mathcal S\setminus\mathcal F_\to\right\}\wedge\\
		&\bigwedge\left\{A^n\to (B^n\circ C^n)\:|\:A\to (B\circ C)\in \mathcal S\setminus\mathcal F_\to\right\}\\
		\text{\normalfont Sat}_{\mathcal F_\to}(n) = & \bigwedge\{F^{n-1}_\psi\to (A^n\to B^n)\to C^{n-1}\:|\:\psi = (A\to B)\to C\in\mathcal F_\to\}
	\end{align*}
\end{definition}

\begin{example}
	With the previous encoding for double negation elimination $\varphi = ((A\to \bot)\to \bot)\to A$ we have
	$$\hspace*{-.3cm}
	\varphi^Q = \exists A^0\forall F^0\left(\neg A^0\wedge\left(F^0\to \exists A^1\left((A^0\to A^1)\wedge(F_0\to (A^1\to \bot)\to \bot)\right)\right)\right),
	$$
	which is a satisfiable QBF since $A^0 = 0, A^1 = 1$ satisfies it for any $F^0$.
\end{example}

\begin{lemma}
	$\varphi$ is not intuitionistically valid iff $\varphi^Q$ is a satisfiable QBF.
\end{lemma}
A proof can be found in the appendix as Lemma~\ref{lemma:QBF}.

A simple counting argument shows that for each $n < |\mathcal F_\to|$ we have $|\text{Valid}(n)|\in \mathcal O(|\mathcal F_\to|^2)$, $|\text{Persistent}(n)|, |\text{Sat}_{\mathcal S}(n)|\in \mathcal O(|\varphi|)$ and $|\text{Sat}_{\mathcal F_\to}(n)|\in \mathcal O(|\mathcal F_\to|)$ so overall we get:
\begin{lemma}
	The size of $\varphi^Q$ is in $\mathcal O(|\varphi|\cdot|\mathcal F_\to| + |\mathcal F_\to|^3)$.
\end{lemma}

@@

First of all recall, that due to Lemma~\ref{equivalid} it is sufficient to consider sentences of the form $\bigwedge\mathcal S\to P$ where all formulas in $\mathcal S$ are of one of the forms
\[\begin{matrix}
	\forall \vec z(A(\vec a)\to (B(\vec b)\wedge C(\vec c)))\indent \forall \vec z((A(\vec a)\wedge B(\vec b))\to C(\vec c))\indent \forall \vec z(A(\vec a)\to (B(\vec b)\vee C(\vec c)))\\
	\forall \vec z((A(\vec a)\vee B(\vec b))\to C(\vec c))\indent \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c))\indent\forall \vec z(\forall xA(\vec a)\to B(\vec b))\\ \forall \vec z(A(\vec a)\to\forall xB(b)), \forall \vec z(\exists xA(\vec a)\to B(\vec b)), \forall \vec z(A(\vec a)\to\exists xB(b)).
\end{matrix}\]
The complete result that we will show is as follows:
\begin{theorem}\label{fullFOtranslation}
	Let $\mathcal S$ be as above and $\mathcal F_\to\subseteq\mathcal S$ contain the formulas of the form $\forall \vec t((A(\vec z_A)\to B(\vec z_B))\to C(\vec z_C))$ and $\mathcal F_\forall\subseteq\mathcal S$ the formulas of the form $\forall \vec z(\forall xA(\vec z_A)\to B(\vec z_B))$. Let $\mathcal F = \mathcal F_\to\cup\mathcal F_\forall$. For every $n$-ary atomic relation $R$ define a new $n+1$-ary relation $R^\#$. For every $\psi\in\mathcal F$ define a new function symbol $f_\psi$. Consider a new binary relation symbol $E$ and define $\vec E(\vec x, u) := \bigwedge\{E(x, u)\:|\:x\in\vec x\}$. Obtain $\mathcal S^\#$ by including formulas as follows:
	\begin{itemize}
		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, u)\to (B^\#(\vec b, u)\circ C^\#(\vec c, u))))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z(A(\vec a)\to (B(\vec b)\circ C(\vec c)))\in\mathcal S$
		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, u)\circ B^\#(\vec b, u))\to C^\#(\vec c, u))$\\for each $\circ\in\{\wedge, \vee\}, \psi = \forall \vec z((A(\vec a)\circ B(\vec b))\to C(\vec c))\in\mathcal S$
		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to(A^\#(\vec a, f(\vec z, u))\to B^\#(\vec b, f(\vec z, u)))\to C^\#(\vec c, u))$ for each $\psi = \forall \vec z((A(\vec a)\to B(\vec b))\to C(\vec c))\in\mathcal S$
		\item  $\forall \vec z\forall u(\vec E(\vec z, u)\to \forall x(E(x, f(\vec z, u))\to A^\#(\vec a, f(\vec z, u)))\to B^\#(\vec b, u))$\\for each $\forall \vec z(\forall xA(\vec a)\to B(\vec t_B))\in\mathcal S$.
		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to A^\#(\vec a, u)\to \forall x(E(x, u)\to B^\#(\vec b, u)))$\\for each $\forall \vec t(A(\vec a)\to \forall xB(\vec b))\in\mathcal S$.
		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to \exists x(E(x, u)\wedge A^\#(\vec z, u))\to B^\#(\vec b))$\\for each $\forall \vec z(\exists xA(\vec a)\to B(\vec b))\in\mathcal S$
		\item $\forall \vec z\forall u(\vec E(\vec z, u)\to A^\#(\vec b)\to \exists x(E(x, u)\wedge B^\#(\vec b, u)))$\\for each $\forall \vec z(A(\vec a)\to \exists xB(\vec b))\in\mathcal S$
	\end{itemize}
	Then $\bigwedge S\to P$ is intuitionistically iff $\exists xE(x, s)\to K\to \bigwedge\mathcal S^\#\to P^\#(s)$ is classically valid, where $s$ is a new constant symbol, $\Sigma$ is the original signature and
	\begin{align*}
		K := &\:\bigwedge\{\forall x\forall u((E(x, u)\to E(x, f_\psi(u)))\:|\:\psi\in\mathcal F\}\wedge\\
		&\:\forall u\left(\exists xE(x, u)\to \bigwedge\{\forall\vec z(\vec E(\vec z, u)\to E(f(\vec z), u))\:|\:\text{$f$ is a function symbol in $\Sigma$}\}\right)\wedge\\
		&\:\bigwedge\{\forall\vec z_1\forall \vec z_2\forall u(A^\#(\vec z_1, u)\to A^\#(\vec z_1, f_\psi(\vec z_2, u)))\:|\:\text{$\psi\in\mathcal F$, $A$ is a relation symbol in $\Sigma$}\}.
	\end{align*}
	The size of the obtained formula is linear in the input. However a $|\mathcal F| + 1$ new function symbols and the new binary predicate $E$ were introduced and each $n$-ary relation symbol has been extended to a $n+1$-ary relation symbol.
\end{theorem}

It is not possible to completely eliminate worlds from the domain as in the propositional case, since counter-models can possibly be infinite (as we will see in more detail later).

@@

As in the propositional case we must encode the Kripke Semantics, this time not only the relation $\preceq$, but also $E$ has to be axiomatized. For that we use the following predicates:
\begin{itemize}
	\item PartialOrder($\preceq$) encoding that $\preceq$ is a partial order.
	\item DomainSubset$(u, w)$ encoding that the domain at $u$ is a subset of that at $w$.
	\item World($u$) encoding that $u$ represents a world.
	\item DomainClosed($u$) which encodes that at $u$ the domain is closed under functions, e.g. contains all interpretations of constants.
	\item Persistent$(u, w)$ which encodes that persistency is satisfied between $u$ and $w$, i.e. for all predicates $A$ and domain elements $x$ with $A(x)$ at $u$ we have $A(x)$ at $w$.
\end{itemize}
Then, analogously to the propositional case, we define:
\begin{align*}
	K(\varphi) = \:& \text{PartialOrder}(\preceq) \wedge \forall u \forall w (u\preceq w\to \text{DomainSubset}(u, w)) \wedge\\
				 & \forall u(\text{World(u)}\to \text{DomainClosed}(u))\wedge \forall u\forall w (u\preceq w\to \text{Persistent}(u, w)).
\end{align*}
Example encodings of these formulas are as follows:
\begin{align*}
	\text{PartialOrder}(\preceq) = &\:\forall u(u\preceq u)\wedge\forall u\forall w(u\preceq w\to w\preceq u\to u = w)\wedge\\&\:\forall u\forall v\forall w(u\preceq v\to v\preceq w\to u\preceq w)\\
	\text{DomainSubset}(u, w) = &\:\forall z(E(z, u)\to E(z, w))\\
	\text{World}(u) = &\:\exists xE(x, u)\\
	\text{DomainClosed}(u) = &\:\bigwedge\{\forall\vec z(\vec E(\vec z, u)\to E(f(\vec z), u))\:|\:\text{$f\in\Sigma$ is a function symbol}\}\\
	\text{Pesistent}(u, w) = &\:\bigwedge\{\forall\vec z(\vec A^\#(\vec z, u)\to A^\#(\vec z, w))\:|\:\text{$A\in\Sigma$ is a relation symbol}\}
\end{align*}

@@

Then $\varphi = \bigwedge\mathcal S\to P$ is intuitionistically valid if and only if
$$\varphi^\circ= \text{World}(s)\to K(\varphi)\to \bigwedge S^\circ\to P^\#(s)$$
is classically valid, where $s$ is a new constant symbol. This was obvious in the propositional case but here it is more nuanced. A proof is given in the appendix as Lemma~\ref{lemma:fo-simplification}.

@@

As in the predicate case we now show the existence of a canonical counter-model (in case such a counter-model exists). Denote the set of formulas $\psi\in\mathcal S$ of the form $\forall\vec z((A(\vec a)\to B(\vec b))\to C(\vec c))$ with $\mathcal F_\to$ and of the form $\forall \vec z(\forall xA(\vec a)\to B(\vec b))$ with $\mathcal F_\forall$. Let $\mathcal F = \mathcal F_\to\cup\mathcal F_\forall$. For each  $\psi\in\mathcal F$ consider a new function symbol $f_\psi$. As before we may assume that $s$ is a least element and we can avoid the quantification over $w$ by considering the formulas
\begin{itemize}
	\item $\forall \vec z\forall u(\vec E(\vec z, u)\to (A^\#(\vec a, f_\psi(\vec z, u))\to B^\#(\vec b, f_\psi(\vec z, u)))\to C^\#(\vec c, u))$\\instead of $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to A^\#(\vec a, w)\to B^\#(\vec b, w))\to C^\#(\vec c, u))$,
	\item $\forall \vec z\forall u(\vec E(\vec z, u)\to (\forall x(E(x, f_\psi(\vec z, u))\to A^\#(\vec a, f_\psi(\vec z, u))))\to B^\#(\vec b, u))$\\instead of $\forall \vec z\forall u(s\preceq u\to\vec E(\vec z, u)\to \forall w(u\preceq w\to \forall x(E(x, w)\to A^\#(\vec a, w)))\to B^\#(\vec b, u))$,
\end{itemize}
and by removing removing the $s\preceq u$ for the remaining formulas in $\mathcal S^\circ$.
We denote the resulting set of formulas by $\mathcal S^\#$.
Then, the previous formulas are equivalid to
$$\varphi^\circ= \forall u(s\preceq u)\to \bigwedge\{\forall \vec t\forall u(u\preceq f(\vec z, u))\:|\:\psi\in\mathcal F\}\to K(\varphi)\to\bigwedge\mathcal S^\#\to P^\#(s).$$
\todo{we already define $\varphi^\circ$, shouldn't this be some $\varphi^\dagger$? Also the phrase before should be: Then,$\varphi^\circ$ is equivalid to ...?}
As in the previous section we can define a tree counter-model for this.

\begin{definition}
	Suppose we are given a counter-model $\mathcal M = (M, I)$ for $\varphi^\circ$. \todo{shouldn't this be $\varphi^\dagger$?} Define $\mathcal M_T = (M_T, I_T)$ as follows:
	\begin{itemize}
		\item $M_T$ are the sequences on $\{(\vec z, \psi)\:|\:\psi\in \mathcal F, \vec z\in M^n\}$.
		\item We interpret $\preceq$ as the prefix-order.
		\item We set $$f_\psi^{I_T}(\vec z, (\vec z_1, \psi_1)\dots (\vec z_n, \psi_n)) = (\vec z_1, \psi_1)\dots (\vec z_n, \psi_n)(\vec z, \psi).$$
		\item For propositional variables $P$, we set $$P^{I_T}((\vec z_1, \psi_1)\dots (\vec z_n, \psi_n)) = P^I(f_{\psi_n}(\vec z_n, \dots(f_{\psi_1}(\vec z_1, s^I))\dots)).$$
	\end{itemize}
\end{definition}

\begin{lemma}
	If $\mathcal M$ is a counter-model to $\varphi^\circ$ \todo{shouldn't this be $\varphi^\dagger$?} then so is $\mathcal M_T$.
\end{lemma}

Even if we don't have the finiteness property of the propositional case, we have still managed to reduce the complexity of possible counter-models because we know that considering worlds of the form $$f_{\psi_1}(\vec z_1, f_{\psi_2}(\vec z_2, \dots f_{\psi_n}(\vec z_n, s)\dots))$$ is sufficient. 
In particular, this allows us to completely eliminate $\preceq$ from the formula, that is we remove the PartialOrder$(\preceq)$
from $K(\varphi)$ and replace $\forall u\forall w\forall x((E(x, u)\wedge u\preceq w)\to E(x, w))$ with $\bigwedge\{\forall \vec z\forall u(E(x, u)\to E(x, f_\psi(\vec z, u)))\:|\:\psi\in\mathcal F_\to\}$ and analogously for persistency to obtain $K$. We set
